# Download RNA-Seq data from GEO {#ch:download}

We will provide an example on how to download data from GEO using the Leucegene CBF-AML RNS-Seq data uploaded at https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE49642. 
The first step in dowloading data from GEO is to download the `SRA` files. For that we need to get `SRX` entries which each one corresponds to a sample in the RNA-Seq cohort.

## Get SRX sample names 

```{r message=FALSE,warning=FALSE}
library(GEOquery)
library(tidyverse)
library(knitr)
library(stringr)
```

Below is an example using one accession number from the [Leucegene data](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE49642).

```{r eval=FALSE}
# Get matrix files for every accession number
series_matrix_info <- function(gse){
gsed <- getGEO(gse,GSEMatrix=TRUE)
gse.mat <- pData(phenoData(gsed[[1]]))
reduced <- gse.mat[,c("title","geo_accession","relation.1")]
write.csv(reduced,file.path("../data",paste(gse,"_",nrow(gse.mat),".csv",sep="")),row.names = FALSE)
}

series_matrix_info("GSE49642") # 43 samples
```


Every row in Table \@ref(tab:SRX) contains sample names (`title`) and `GSM` accession numbers. In order to download a particular sample we need the `SRA` terms which are the names starting with: `SRX***` in the `relation.1` column. The structure of the matrix might change across different studies but you should be able to find `SRX` entries hidden somewhere in the `GSEMatrix`!

```{r SRX,message=FALSE}
matrix_file <- list.files(path = file.path("data"),pattern = "GSE",full.names = TRUE)
GSEmatrix <- read_csv(matrix_file)

kable(GSEmatrix[1:5,],caption="SRX sample names.")
```

With some string processing we can extract the `SRX` entries. 

```{r message=FALSE}
GSEmatrix$SRX <- stringr::str_extract(string = GSEmatrix$relation.1,pattern = "SRX[0-9][0-9][0-9][0-9][0-9][0-9]")
GSEmatrix$relation.1 <- NULL
kable(head(GSEmatrix))
```

## Create NCBI query

```{r}
search_ncbi <- paste(GSEmatrix$SRX,collapse=" OR ")
search_ncbi
```

Paste the search `r search_ncbi` into NCBI https://www.ncbi.nlm.nih.gov/sra and follow the intructions in https://www.ncbi.nlm.nih.gov/sra/docs/sradownload/#download-sequence-data-files-usi **Download sequence data files using SRA Toolkit** to download all the `SRR` run names and information of the runs. 

```{bash eval=FALSE}
# Files are saved in the home directory under ncbi/public/sra
prefetch --option-file SraAccList_CBF-AML_Leucegene.txt
```

SRA files can then be converted to `fastq` files with `fastq-dump --split-files`. 
